---
title: "Coursera Repro Res Wk2 Project"
author: "Nick Paschalidis"
date: "22/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = FALSE)
```

## 1. Loading and the data

Code to download, unzip and read the data.
A master file called `rawdata` is created containing all the activity data

```{r load data}
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", destfile = "rawdata_activity.zip")

unzip("rawdata_activity.zip")

rawdata <-read.csv("activity.csv", header=TRUE)
```

## 2. What is mean total number of steps taken per day?

Code to calculate  mean total number of steps per day and plot.

```{r mean average daily steps} 
Sortbydate <- tapply(rawdata$steps, rawdata$date, sum)
hist(Sortbydate, xlab ="no. of steps", main="Daily steps", breaks = 30)
mean(Sortbydate, na.rm = TRUE)
median(Sortbydate, na.rm = TRUE)
```

## 3. What is average daily activity pattern?

Code to calculate  average total number of steps per day and plot.

```{r daily pattern}
Interval_steps <- tapply(rawdata$steps, rawdata$interval, mean, na.rm = TRUE)
plot(Interval_steps, xlab = "interval", ylab = "No of Steps", main = "Average total number of steps")
```

The 5-minute interval tha contains the maximum number of steps is identified 
using the following code.

```{r intervan with max activity}
Sort_intervals <- names(sort(Interval_steps, decreasing =TRUE) [1])
Sort_steps <- sort(Interval_steps, decreasing =TRUE) [1]
```

Interval 835 is associated with maximum activity, ~206 steps. 

## 4. Imputing missing Values

Calculating missing values before imputing
```{r NAs before imputing}
str(rawdata)
missing_values <- sum(is.na(rawdata$steps))
missing_values
```


```{r imputing NAs}
Interval_steps <- tapply(rawdata$steps, rawdata$interval, mean, na.rm = TRUE)
rawdata.split <- split(rawdata, rawdata$interval)
for(i in 1:length(rawdata.split)){
  rawdata.split[[i]]$steps[is.na(rawdata.split[[i]]$steps)] <- Interval_steps[i]
}
rawdata2 <- do.call("rbind", rawdata.split)
rawdata2 <- rawdata2[order(rawdata2$date) ,]
```

Missing values after imputing
 
```{r NAs after imputing}
missing_values <- sum(is.na(rawdata2$steps))
str(rawdata2)
missing_values
```


## 5. Are there differences in activity patterns between weekdays and weekends?

Code to compute and present activity pattern for weekends and weekdays

```{r}
rawdata2$day <- ifelse(weekdays(as.Date(rawdata2$date)) == "Saturday" | weekdays(as.Date(rawdata2$date)) == "Sunday", "weekend", "weekday")

# Average steps per interval for weekends and weekdays
Interval_steps_wkend <- tapply(rawdata2[rawdata2$day == "weekend" ,]$steps, rawdata2[rawdata2$day == "weekend" ,]$interval, mean,na.rm=TRUE)
Interval_steps_wkdays <- tapply(rawdata2[rawdata2$day == "weekday" ,]$steps, rawdata2[rawdata2$day == "weekday" ,]$interval, mean, na.rm=TRUE)

# Plot Weekday and weekend activity
par(mfrow=c(1,2))
plot(as.numeric(names(Interval_steps_wkdays)), 
     Interval_steps_wkdays, 
     xlab = "Interval", 
     ylab = "No. of Steps", 
     main = "Week day activity")
plot(as.numeric(names(Interval_steps_wkdays)), 
     Interval_steps_wkdays, 
     xlab = "Interval", 
     ylab = "No of. Steps", 
     main = "Weekend activity")
```

